An Improved Method for Coupling Hydrodynamics with Astrophysical Reaction Networks

### 2021-12-29 simplified_sdc_intro/paper.tex
more work

%\documentclass[times,modern]{aastex63}
\documentclass[times,preprint]{aastex63}

@@

@@

We use a network with 28 nuclei and 96 rates that captures He burning
with links for $(\alpha,p)(p,\gamma)$ reactions and the rates
involving \isot{N}{14} to bypass
$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$ as described in
\cite{shenbildsten}.  The rates are taken from the ReacLib library
\citep{reaclib} and the network is written by
\pynucastro~\citep{pynucastro} directly in the C++ format that our
code requires.  The nuclei and links in the network are shown in
Figure~\ref{fig:subch_network}.  This moderate-sized network with many
reaction pathways (forward and reverse) make it a good test for the
coupling between hydro and reactions.  An overview of the network is
shown in Figure~\ref{fig:subch_network}.

@@

Nonaka plots

@@

We have not focused on performance or load-balancing the reactions in
the tests that we ran, but these limited runs show that the
simplified-SDC method runs as fast (or faster) than the Strang method
for the same tolerances and CFL, despite having to do the advection
twice.  We would expect similar behavior for any simulations that are
dominated by the cost of reactions.

@@

We presented a simplified spectral deferred corrections scheme for coupling
hydrodynamics and reactions.

Future work is to extend this methodology to MHD.

>>>>

\documentclass[times,modern]{aastex63}
\documentclass[times,preprint]{aastex63}

@@

For both Strang and simplified-SDC, we will use a numerical Jacobian.
VODE computes it internally for Strang and we use the method from
Appendix \ref{app:jacobian} for the simplified-SDC method.  VODE
employs Jacobian-caching, so the Jacobian does not need to be
re-evaluated each time it is used.

@@

We use a network with 28 nuclei and 96 rates that captures He burning \MarginPar{should I do a few aprox13 runs for comparison?}
with links for $(\alpha,p)(p,\gamma)$ reactions and the rates
involving \isot{N}{14} to bypass
$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$ as described in
\cite{shenbildsten}.  The rates are taken from the ReacLib library
\citep{reaclib} and the network is written by
\pynucastro~\citep{pynucastro} directly in the C++ format that our
code requires.  The nuclei and links in the network are shown in
Figure~\ref{fig:subch_network}.  This moderate-sized network with many
reaction pathways (forward and reverse) make it a good test for the
coupling between hydro and reactions.  An overview of the network is
shown in Figure~\ref{fig:subch_network}.  This network is available 
as {\tt subch2} in \citep{microphysics}.

@@

We now look at the evolution of the state while burning by storing the
thermodynamic state internal to VODE each substep as it evaluates the
righthand side of the ODE system it is integrating.  We pick a point
on the symmetry axis a few zones behind the inward moving shock.
Figure~\ref{fig:nonaka_plot} shows the evolution of \isot{Ni}{56} over
a single coarse-grid timestep (4 fine-grid timesteps).

@@

@@

We presented a simplified spectral deferred corrections scheme for coupling
hydrodynamics and reactions.  We demonstrated that for a moderate-sized reaction
network, the simplified-SDC method

We have not focused on performance or load-balancing the reactions in
the tests that we ran, but these limited runs show that the
simplified-SDC method runs as fast (or faster) than the Strang method
for the same tolerances and CFL, despite having to do the advection
twice.  We would expect similar behavior for any simulations that are
dominated by the cost of reactions.  If the results above a robust across
different problems and networks, then this suggests that we can run 
the SDC simulation with larger CFL and/or less strict reaction tolerances,
which would translate into a large performance boost.

This new time-integration method is freely available in \castro, and
we will continue to explore it on other problems.  We are particularly
interested in massive star evolution with this method, where the
vigorous shell burning leading up to core-collapse would be an ideal
application for the simplified-SDC method.  In a follow-on paper, we
will show how to couple this new integration method to a table for
evaluating an NSE distribution in the iron core of a massive star.

Future work is to extend this methodology to MHD.  This is very
straightforward since the simplified-SDC method operates solely on the
thermodynamic state.  We will also explore how to couple radiation
hydrodynamics, where we need to do an implicit update for the
radiation energy.
